### 2024-03-10 outputs/paper/paper.qmd
download data script

---
title: ""
title-block-banner: true
abstract: ""
thanks: "Code and data supporting this analysis are available at https://github.com/TEJMaster/Replication-For-Money-Illusion.git Replication on Social Science Reproduction platform available at: https://www.socialsciencereproduction.org/reproductions/1671/published/index"
author: "Terry Tu, Jingyi Shen, Yaning Jin "
date: today
date-format: long
format: pdf
toc: true
toc-depth: 1
number-sections: true
bibliography: references.bib
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---