### 2024-03-11 outputs/paper/paper.qmd
model

---
title: ""
title-block-banner: true
abstract: ""
thanks: "Code and data supporting this analysis are available at "
author: "Terry Tu, Jingyi Shen, Yaning Jin "
date: today
date-format: long
format: pdf
toc: true
toc-depth: 1
number-sections: true
bibliography: references.bib
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE, message=FALSE}
## Import necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(readr)
```

## Introduction

## Data

```{r include=FALSE, message=FALSE}
# read the data file and store it in the 'data' variable
data <- read_csv(here("outputs", "data", "cleaned_building_data.csv"))
```

## Model

## Result

## Discussion