### 2024-03-11 outputs/paper/paper.qmd
added new scripts

---
title: "Building Integrity in Retrospect: Assessing the Impact of Age on 2023 Safety Evaluations"
title-block-banner: true
abstract: ""
thanks: "Code and data supporting this analysis are available at "
author: "Terry Tu, Jingyi Shen, Yaning Jin "
date: today
date-format: long
format: pdf
toc: true
toc-depth: 1
number-sections: true
bibliography: references.bib
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE, message=FALSE}
## Import necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(readr)
```

## Introduction

## Data

```{r include=FALSE, message=FALSE}
# read the data file and store it in the 'data' variable
cleaned_data <- read_csv(here("outputs", "data", "cleaned_building_data.csv"))
```

## Model

## Result

```{r echo=FALSE}
# Construct a linear model
lm_model <- lm(CURRENT.BUILDING.EVAL.SCORE ~ AGE, data = cleaned_data)

# Summary of the linear model
summary(lm_model)
```
```{r echo=FALSE}
# Create a scatter plot with a regression line to visualize the relationship
improved_plot <- ggplot(cleaned_data, aes(x = AGE, y = CURRENT.BUILDING.EVAL.SCORE)) +
  geom_point(alpha = 0.5, size = 1.5) +  # Adjust transparency and size of points
  geom_smooth(method = "lm", color = "blue", size = 1) +  # Change the size of the regression line
  labs(
    x = "Age of Building",
    y = "Current Building Evaluation Score",
    title = "Relationship between Building Age and Evaluation Score"
  ) +
  theme_minimal() +  # Use a minimal theme
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.position = "none"  # Hide the legend if not necessary
  )

# Print the improved plot
print(improved_plot)
```

## Discussion